Microlensing as a possible probe of the internal structure of quasars

### 2015-07-10 paper1/fitting.tex
editing

@@

\begin{figure}
\centering
\includegraphics[width=0.9\hsize]{plots/burnin_cg.eps}
\caption{\label{fig:datafitting} The $\chi^2$ of each step of the full
  MCMC for one case.}
\end{figure}

@@

In this section, we attempt to mock three datasets, each for crescent, uniform-disk and Gaussian disk case for a given fiducial parameters and contaminate it with systematic and random noise, and fit \& recover the model parameters. Broadly we have 9 cases to discuss, we name these nine cases as different permutations (couplet) of three letters C, D and G for Crescent, Uniform-Disk and Gaussian Disk case respectively. The first letter of the couplet represents the fit model and the second represents the dataset. For example, CG represents the case where we fit Gaussian disk data to a crescent model. Similarly, DC respresents the case when crescent data is being fit with uniform-disk model. We explored the parameter space and carried out a likelihood analysis on all three parameter sets using Markov-chain Monte-Carlo (MCMC).

\subsection{Mock datasets and model parameters}
The datasets for each source model is in the form of light curves, the brightness of the source observed at different time. Figure \ref{fig:datafitting} shows the three datasets. These datasets are generated using the magnification map shown in figure \ref{fig:magnification_map} if the source of the respective model moves from point C to B. The data assumed was very  ideal, including 250 different data points at regular interval of time. We added Gaussian random noise to this dataset of width 10\% of the brightness at the respective time, which makes the signal to noise of about $\sim$150. After adding random noise, the three datasets are looking highly contaminated and no clear intuitive behaviour of fit can be seen.

The fiducial model assumed to generate these mock datasets are:
\begin{equation}
	[R_p, R_n, a, b] = [50.0, 30.0, 15.0, 10.0]  \ \ \ \rm{for\ Crescent\ model},
	\label{eqn:cp}
\end{equation}
\\
where, $R_p$ and $R_n$ are the radius of outer and inner circles of the crescent respectively and $a$ and $b$ are the coordinates of the inner circle center with center of outer circle as the origin.

\begin{equation}
	R_p = 50.0 \ \ \ \rm{for\ Uniform\ disc\ model}
\end{equation}

\begin{equation}
	R_p = 50.0 \ \ \ \rm{for\ Gaussian\ disk\ model}
\end{equation}
\\
where, $R_p$ being the size of the uniform disc and Gaussian disc source model. We also have some nuisance parameters in all these cases: the beginning of the event (I), the end of the event (E), normalisation of the lightcurve (N). We marginalised over all these nuisance parameters. When we kept these three nuisance parameters free to vary in MCMC, it adapts to the amount of data for which the maximum likelihood is most optimal to find out. We talk more about this adaptaion in the next section. Secondly, We try to fit the these datasets for a crossing of the source from a different caustic, in this case, from point A to B in figure \ref{fig:magnification_map}. This adds a natural systematic in the dataset and the data is no more ideal. With these three characterstics -- random noise, systematic noise and nuisance parameters -- we make the dataset very general and highly contaminated to mock the real observations.

The aim of this exercise was to explore the degeneracies in  various model parameters and recover the correct fiducial model assumed from a highly contaminated dataset. Also, if it is possible that one can simply distinguish between circularly symmetric and asymmetric sources or uniform and Gaussian sources on the basis of maximum likelihoods.

>>>>

In this section we simulate the fitting of source models to
lightcurves in the presence of random and systematic errors.  We
generate mock data for three source models --- a crescent, a uniform
disc and a Gaussian disc --- and then fit each mock light curve to all
three source models.  We denote the possible cases letter pairs, with
the first letter denoting the model and the second letter representing
the source: thus CG means a {\em C\/}rescent model was fitted to mock
data from a {\em G\/}aussian source, DC stands for a uniform {\em
  D\/}isc model fitted to mock data from a {\em C\/}rescent source,
and so on.

@@

\subsection{Mock datasets}

Mock light curves were generated by running a source across a
caustic, specifically along the line from point C to point B in
Figure~\ref{fig:magnification_map}.  The model light curves, to which
these are fitted, are made by running a source across a different
caustic, along the line from point A to point B in the same figure.
That is, the mock data are generated with a clean but not ideal fold
caustic and then fitted with another such caustic.  This mimics the
unavoidable systematic error of not knowing the caustic exactly.

The crescent source had parameters (as explained in
\S\ref{subsec:crescent}) $R_p$ and $R_n$ being the radii of the outer
and inner circles and $(a,b)$ being the coordinates of the inner
circle with respect to the centre of the outer circle.  The parameter
values were
\begin{equation}
   (R_p, R_n, a, b) = (50.0, 30.0, 15.0, 10.0).
\label{eqn:cp}
\end{equation}
The uniform disc had the same (outer) radius as the crescent.  In the
Gaussian source, we set $3\sigma=50$.  For convenience, below we will
refer to $R_p$ of the Gaussian source, by which we actually mean
$3\sigma$.

Each mock light curve also has three nuisance parameters, namely the
beginning and end of the event and the brightness normalization.
These are to be marginalised out by the MCMC.

Figure~\ref{fig:datafitting} shows the three light curves.  Each has
250 points regularly spaced in time, with Gaussian noise at the level
of 10\% of the current brightness.  The effective signal to noise can
be considered to be $0.1\times\sqrt{250}\simeq150$.

\subsection{Fitting}

Parameter fitting and marginalisation was done by Markov-chain Monte
Carlo (MCMC).  Figure~\ref{fig:burnin} shows the progress of reduction
of $\chi^2$.

\begin{figure}
\centering
\includegraphics[width=0.9\hsize]{plots/burnin_cg.eps}
\caption{\label{fig:burnin} The $\chi^2$ of each step of the full MCMC
  for one case.}
\end{figure}

@@
